The Name of the Title is Hope

### 2019-03-26 emsoft-2019/main.tex
background subsections, and other fix to several formulas

\section{Introduction}
\section{Background}
\section{Problem Description}
In this work we present an innovative method for the off-line design of robust MPC controllers (explicit), for linear time invariant systems.

To guarantee stability of the system at runtime, at design time the designer specifies the maximal disturbance value the system can tolerate: we call this numeric value \texttt{delta}. This delta has to be an upper bound for both errors generated by the system itself (e.g. noise from sensors) and disturbance coming from sources external to the system (e.g. friction).

The system is then designed to be robust against a maximal disturbance of value \texttt{delta}.

Among the sources of disturbance targeting the controller, our goal is to assure that the approximation error caused by the use of finite precision arithmetic is also bounded by this design value \texttt{delta}.

@@

(ii) Second, we want to take advantage of this disturbance delta, and reduce the arithmetic precision used for computations and storage, while still guarantee that the approximation error generated is bounded by delta. Our intuition is that a greater value for delta allows for a reduced precision for computations (e.g. with respect to 32bits).

Given an initial delta, we design the controller by solving an off-line optimization problem. We rely on MATLAB multi-parametric toolbox to solve such optimization problem~\cite{matlabMPT, matlabYALMIP}. 
The output of the suite is a continuous piecewise affine (PWA) function, together with a partitioning of the domain of \statevarmath. Each element in this partitioning is called a region.
A generic region $i$ consists in a n-polytope, uniquely identified by a set of inequalities $H_{i}\statevar<=K_{i}$.

@@

Each region $i$ is associated with an activation function $u_{i,k}=F_{i}x_{k}+G_{i}$.

@@

A drawback in the design of explicit MPC is that both regions boundaries (H and K) and activation functions (F and G), have to be stored on the micro-controller. Usually these devices have limited memory in the order of KBs.

At runtime, the value of state variable \statevarmath  is compared against the polytopes bounds stored on the device: depending on the region containing \statevarmath, the corresponding activation function is computed.

@@

\texttt{err} generates instability when it is time to verify which region \qstatevarmath belongs to: given the value of \statevarmath the region should be $i$, but because of \texttt{err}, actually region $j$ is selected. This scenario becomes more intuitive when \qstatevarmath falls \texttt{close} to the border between two generic regions $i$ and $j$, but in general it depends on the magnitude of \texttt{err}.

@@

Our goal is to build a feasible geometrical space around the border between regions $i$ and $j$, where actually it is feasible that \statevarmath belongs to region $i$ but \qstatevarmath follows in region $j$.

We call this geometrical space the \texttt{tube}.

@@

The total size of the \texttt{tube} is then:
\begin{equation}\label{eq:epsilontot}
\varepsilon=\varepsilon_{A/D}+\varepsilon_{Q}
\end{equation}
\section{Finite Precision Implementation}
The output of the controller can be affected by two main errors: (i) the controller chooses the wrong activation function because of $\varepsilon$ in (\ref{eq:epsilontot}), while (ii) is the approximation error deriving from the finite precision arithmetic used to compute the activation function itself~\cite{imperialrmpc}.

The effect of selecting the wrong activation function are similar to instability in the switch reported in Listing \ref{lst:caseof}. Suppose an hypothetic scenario where all the measurements were done in infinite precision and without any uncertainty: the controller $i$ would be activated.

Instead, because of these limitations controller $j$ is selected. The resulting error is then the difference between the output of the two branches $i$ and $j$.

The system has to be robust against a mistake in choosing controller $i$ instead of $j$ (or vice-versa) only when \qstatevarmath falls into the \texttt{tube} between $i$ and $j$, and not for all the points in the two regions.

In the following, we assume $i$ and $j$ are the index of two generic regions in \statespace:
\begin{flalign}
\label{eq:maximization}
&\max_{\forall i,j\;|\;neighbour(i,j)}|u_{i}-u_{j}| = \\
&\max_{\forall i,j\;|\;neighbour(i,j)}|F_{i}\statevar+G_{i} - (F_{j}\statevar+G_{j})|\nonumber
\end{flalign}

Because of the linearity of the function $u_{i}-u_{j}$, and because of the convexity of the regions $i$, $j$ it is enough to evaluate function (\ref{eq:maximization}) at the corners of the \texttt{tube}.

@@

Starting from the equation of $border_{i,j}$, we delimit the geometrical space where it makes sense to compute (\ref{eq:maximization}) with: 
\begin{equation}
\begin{aligned}
border_{i,j} >= -\varepsilon\\
border_{i,j} <= \varepsilon
\end{aligned} 
\end{equation}
where $\varepsilon$ is defined in (\ref{eq:epsilontot}).

We call such geometrical space the \texttt{tube} between $i$ and $j$.

When \qstatevarmath belongs to the \texttt{tube}, it might be that $u_{i}$ is activated instead of $u_{j}$ (or vice-versa). Otherwise, when \qstatevarmath does not belong to the tube, no matter the error \texttt{err}, the right activation function is activated, and we consider only the error deriving from the computation itself.

@@

In (7) the domain of \qstatevarmath are all the points in region $i$, but also all the values in the \texttt{tube} between region $i$ and any of the neighbors of $i$. Even for values that do not belong to region $i$ it might be the case that $u_{i}$ is (erroneously) activated. In particular, this is true for values in the \texttt{tube}.

We compute (\ref{eq:fperror}) for all regions in \statespace\space and we take the maximum value:

@@

The disturbance \texttt{delta} then has to be an upper bound to the summation of (\ref{eq:maximization}) and (\ref{eq:fperror}).

@@

\section{Precision Tuning}
In formula (\ref{eq:epsilontot}), the equation for $\varepsilon$ contains: $\varepsilon_{A/D}$ that is the error introduced by the analog-digital conversion, and $\varepsilon_{Q}$ that is generated by the quantization of hyperplanes $H\statevar<=K$ in a finite number of bits.
This $\varepsilon_{Q}$ can be tuned based on memory availability in the micro-controller.
The rule for $\varepsilon_{Q}$ is the following:
\begin{equation}\label{eq:quantizationlines}
\varepsilon_{Q} > (H-(\hat{H})_{p})\qstatevar+(K-(\hat{K})_{p})
\end{equation}
This inequality assures that the distance between any hyperplane represented in infinite precision, and its counter-part quantized in p bits, is bounded by $\varepsilon_{Q}$.

Our goal is to solve (\ref{eq:quantizationlines}) with respect to p: find the minimal-precision p such that the inequality holds.

The main advantage in doing this (compared to fixing the precision a priori), is that the value of p can be tuned based on the memory available in the micro-controller.

To solve (\ref{eq:quantizationlines}) we used Daisy: a static analyzer for finite precision expressions, able to provide a sound upper-bound to the maximal approximation error of an expression with respect to its real (infinite precision) counterpart. Since the magnitude of the round-off error depends on the range of input variables, any variable encoded in Daisy has to be bounded.

Since the state variable \statevarmath is bounded, and F,G,H, and K are vectors of constants, we can rely on Daisy for this verification step.

In a similar way, we use Daisy to solve the inequality in (\ref{eq:delta}) with respect to p.

\begin{flalign}
\label{eq:deltaminusmax}
&delta - \Big(\max_{\forall i,j\;|\;neighbour(i,j)}|u_{i}-u_{j}|\Big)>=\\
& \max_{\forall\;\regionimath{i}\;in\;\statespace} err(u_{i})_{p}\nonumber
\end{flalign}
We want to point out that the left side of the inequality is not parametrized in p. Only the right size of (\ref{eq:deltaminusmax}) actually depends on the precision p. We encode the inequality (\ref{eq:deltaminusmax}) in Daisy, to detect what is the minimal precision that satisfy the inequality.

@@

\begin{lstlisting}[language=Python,numbers=left,numbersep=3pt,frame=lines,keepspaces=true,escapeinside={(*}{*)},caption=design of robust MPC with verification and precision tuning,label={lst:alg}]
 delta=input()
 (*$\varepsilon_{Q}$*)=input()
 assert (delta>=0 and (*$\varepsilon_{Q}$*)>0)
 epsilon=(*$\varepsilon_{A/D}$*)+(*$\varepsilon_{Q}$*) #size of the tubes

 while True:
   design_robust_MPC(delta)
   maxUij = compute (*(\ref{eq:maximization})*) with size(tube)=(*$\varepsilon$*)
   if delta > maxUij:
     max_err=delta-maxUij
     UNI_MIX_precision(F,G,max_err)
     UNI_MIX_precision(H,K,(*$\varepsilon_{Q}$*))
     break
   else:
     delta=maxUij+(*$\varepsilon_{SAFE}$*)

\end{lstlisting}
In Listing \ref{lst:alg} we describe the procedure used to design an MPC controller robust to both (\ref{eq:maximization}) and (\ref{eq:maxfperror}).

@@

Even if it possible to assign value zero to delta~\cite{imperialrmpc}, we do not encourage such initialization: it is going to fail the analysis at least once because the error in (\ref{eq:maxfperror}) is going to be greater than zero (even in case (\ref{eq:maximization}) is zero). A better initialization would be to set delta to an arbitrary small value, slightly greater than zero.

@@

In case the conditional statement is verified,
the precision tuning phase starts.
For the precision tuning of activation functions, we allow an error that is bounded by \texttt{max\_err}. This is because we still need to satisfy (\ref{eq:delta}).
On the other hand, the quantization of polytopes borders (the region bounds in \statespace) can produce an error that is at most $\varepsilon_{Q}$, in this way we satisfy 
(\ref{eq:quantizationlines}).

In case the conditional fails and \texttt{maxUij} is greater than delta, the controller needs to be re-designed with a robustness value that is at least equal to the current value of \texttt{maxUij}. The same consideration done for the initialization of input parameters holds also here: the constant $\varepsilon_{SAFE}$ is used to relax the value delta for the controller, to a value slightly greater than \texttt{maxUij}. In this way, we aim to give some space to the precision tuning phase in the next iteration of the loop. Otherwise, in case of $\varepsilon_{SAFE}$ equal to zero, the next iteration of the precision tuning phase is going to require unnecessarily wide precision values p, usually greater than 32 bits. 
We remark that the point of the analysis is to find a low precision configuration for bounds and activation functions: in case the analysis outputs a precision greater than 32bits, we fail in our goal.
Then, with a minimal alteration to the upper bound of delta, we sensibly reduce the precision needed for $F$ and $G$, with respect to the standard 32 bits precision.

>>>>

\section{Introduction}
\section{Background}
\subsection{Robust MPC}
\subsection{Finite Precision Arithmetic}
\subsection{Daisy}
\section{Problem Description}
In this work we present an innovative method for the off-line design of robust MPC controllers (explicit), for linear time invariant systems.

To guarantee stability of the system at runtime, at design time the designer specifies the maximal disturbance value the system can tolerate: we call this numeric value \texttt{delta}.

\texttt{delta} has to be an upper bound for both errors generated by the system itself (e.g. noise from sensors) and disturbance coming from sources external to the system (e.g. friction).

The system is then designed to be robust against any kind of disturbance up to a maximal value \texttt{delta}.

Among the sources of disturbance targeting the controller, our goal is to assure that the approximation error caused by the use of finite precision arithmetic is also bounded by this design property.

@@

(ii) Second, we want to take advantage of this disturbance delta, and reduce the arithmetic precision used for computations and storage, while still guarantee that the approximation error is bounded by delta. Our intuition is that a greater value for delta allows for a reduced precision for computations (e.g. with respect to 32bits precision).

Given an initial delta, we find the right controller for a system by solving an off-line optimization problem. We rely on MATLAB multi-parametric toolbox to solve such optimization problem~\cite{matlabMPT, matlabYALMIP}. 
The output of the suite is a continuous piecewise affine (PWA) function, together with a partitioning of the domain of \statevarmath. Each element in the partitioning is called a region.
A generic region $i$ consists in a n-polytope, uniquely identified by a set of inequalities in the form: $H_{i}\statevar<=K_{i}$.

@@

@@

Each region $i$ is associated with an activation function $u_{i,k}=F_{i}x_{k}+G_{i}$. At runtime, the value of state variable \statevarmath  is compared against the polytopes bounds: depending on the region $i$ containing \statevarmath, the corresponding activation function $u_{i}$ is computed.

A drawback in the design of explicit MPC is that both regions boundaries (H and K) and activation functions (F and G), have to be stored on the micro-controller. Usually these devices have limited memory, in the order of KBs.

@@

\texttt{err} generates instability when it is time to verify which region \qstatevarmath belongs to: it can happen that \statevarmath satisfies the bounds of region $i$, but because of \texttt{err}, \qstatevarmath belongs to region $j$. This scenario becomes more intuitive when \qstatevarmath falls \texttt{close} to the border between two neighbor regions $i$ and $j$, but in general this instability depends on the magnitude of error \texttt{err}.

@@

Our goal is to build a feasible geometrical space around the border between two neighbor regions $i$ and $j$, where actually it is feasible that \statevarmath belongs to region $i$ but \qstatevarmath follows in region $j$ (or vice-versa).
We call this geometrical space the \texttt{tube}.

@@

The total size of the \texttt{tube} is then:
\begin{equation}\label{eq:epsilontot}
\varepsilon=\varepsilon_{A/D}+\varepsilon_{Q}
\end{equation}
\section{Finite Precision Implementation}
The output of the controller can be affected by two main errors: (i) the controller chooses the wrong activation function because of $\varepsilon$ in (\ref{eq:epsilontot}), and (ii) the approximation error deriving from finite precision arithmetic used to compute the activation function itself~\cite{imperialrmpc}.

The effect of picking the wrong activation function are similar to instability in the switch reported in Listing \ref{lst:caseof}.

In an hypothetic scenario where all the measurements were done in infinite precision arithmetic and without any uncertainty, we assume the controller $i$ would be activated. Instead, because we cannot rely on infinite precision, controller $j$ is selected.

The error committed is the difference between the output of the two branches $i$ and $j$.

The system has to be robust against a mistake in choosing controller $i$ instead of $j$ (or vice-versa), only when \qstatevarmath falls into the \texttt{tube} between $i$ and $j$, and not for all the points in the two regions. Moveover, we are interested in the worst case scenario where is maximized the difference between the activation functions of any two neighbor regions.

In the following, we assume $i$ and $j$ are the index of two generic regions in \statespace:
\begin{flalign}
\label{eq:maximization}
&\max_{\forall i,j\;|\;neighbour(i,j)}|u_{i}-u_{j}| = \\
&\max_{\forall i,j\;|\;neighbour(i,j)}|F_{i}\statevar+G_{i} - (F_{j}\statevar+G_{j})|\nonumber
\end{flalign}
where \statevarmath belongs to the \texttt{tube} between region $i$ and $j$.
Because of the linearity of the function $u_{i}-u_{j}$, and because of the convexity of the regions $i$ and $j$, it is enough to evaluate function (\ref{eq:maximization}) at the corner points of the \texttt{tube}, instead of solving a maximization problem.

@@

Starting from the equation of $border_{i,j}$, we delimit the geometrical space where it makes sense to compute (\ref{eq:maximization}) with: 
\begin{equation}
\begin{aligned}
(border_{i,j} >= -\varepsilon) \land
(border_{i,j} <= \varepsilon)
\end{aligned} 
\end{equation}
where $\varepsilon$ is defined in (\ref{eq:epsilontot}). We call such geometrical space the \texttt{tube} between $i$ and $j$.

When \qstatevarmath belongs to the \texttt{tube}, it might be that $u_{i}$ is activated instead of $u_{j}$ (or vice-versa). Otherwise, when \qstatevarmath does not belong to the tube, no matter the error $\varepsilon$, the right activation function is activated, and we consider only the error deriving from the computation itself.

@@

In (7) the domain of \qstatevarmath are all the points in region $i$, but also all the values in the \texttt{tube} between region $i$ and any of the neighbors of $i$. Even if some points in the tube do not belong to region $i$, it might that $u_{i}$ is (erroneously) activated.

We compute (\ref{eq:fperror}) for all regions in \statespace\space and we take the maximum value of the error (worst case):

@@

The disturbance \texttt{delta}  has to be an upper bound to the summation of (\ref{eq:maximization}) and (\ref{eq:maxfperror}).

@@

\section{Precision Tuning}
In formula (\ref{eq:epsilontot}), the equation for $\varepsilon$ contains: $\varepsilon_{A/D}$ that is the error introduced by the analog-digital conversion, and $\varepsilon_{Q}$ that is generated by the quantization of hyperplanes $H\statevar<=K$ in a finite number of bits.
This $\varepsilon_{Q}$ can be tuned based on memory availability in the micro-controller.
The rule for $\varepsilon_{Q}$ is the following:
\begin{equation}\label{eq:quantizationlines}
\varepsilon_{Q} > (H-\hat{H}_{p})\qstatevar+(K-\hat{K}_{p})
\end{equation}
This inequality assures that the distance between any hyperplane represented in infinite precision (H and K), and its counter-part quantized in p bits ($(\hat{H})_{p}$ and $(\hat{K})_{p}$), is bounded by $\varepsilon_{Q}$.

Our goal is to solve (\ref{eq:quantizationlines}) with respect to p: assign an arbitrary value to $\varepsilon_{Q}$ and find the minimal-precision p such that the inequality holds.

The main advantage in this approach (compared to fixing the precision p a priori), is that the value of p can be tuned based on the memory availability in the micro-controller.

To solve (\ref{eq:quantizationlines}) we used Daisy: a static analyzer for finite precision expressions, ables to provide a sound upper-bound to the maximal approximation error of a formula with respect to its real (infinite precision) counterpart. Since the magnitude of the round-off error depends on the range of input variables, any variable encoded in Daisy has to be bounded.

Since the state variable \statevarmath is bounded by \statespace, and F,G,H, and K are vectors of constants, we can rely on Daisy for this verification step.

In a similar way, we use Daisy to solve the inequality in (\ref{eq:delta}) with respect to p:
\begin{flalign}
\label{eq:deltaminusmax}
&delta - \Big(\max_{\forall i,j\;|\;neighbour(i,j)}|u_{i}-u_{j}|\Big)>=\\
& \max_{\forall\;\regionimath{i}\;in\;\statespace} err(u_{i})_{p}\nonumber
\end{flalign}

Be aware that the left side of the inequality is not parametrized in p. Only the right size of (\ref{eq:deltaminusmax}) actually depends on the precision p. After we come up with a numerical value for the left side of the inequality, we encode (\ref{eq:deltaminusmax}) in Daisy in the same way done for (\ref{eq:quantizationlines}).

@@

\begin{lstlisting}[language=Python,numbers=left,numbersep=3pt,frame=lines,keepspaces=true,escapeinside={(*}{*)},caption=design of robust MPC with verification and precision tuning,label={lst:alg}]
 delta=input()
 (*$\varepsilon_{Q}$*)=input()
 assert (delta>=0 and (*$\varepsilon_{Q}$*)>0)
 (*$\varepsilon$*)=(*$\varepsilon_{A/D}$*)+(*$\varepsilon_{Q}$*)

 while True:
   design_robust_MPC(delta)
   maxUij = compute (*(\ref{eq:maximization})*) with size(tube)=(*$\varepsilon$*)
   if delta > maxUij:
     max_err=delta-maxUij
     UNI_MIX_precision(F,G,max_err)
     UNI_MIX_precision(H,K,(*$\varepsilon_{Q}$*))
     break
   else:
     delta=maxUij+(*$\varepsilon_{SAFE}$*)
\end{lstlisting}
In Listing \ref{lst:alg} we describe the procedure used to design a robust MPC controller such that (\ref{eq:delta}) is respected.

@@

Even if it possible to assign value zero to delta~\cite{imperialrmpc}, we do not encourage such initialization value: it is going to fail the analysis (at least) for the first iteration because the error in (\ref{eq:maxfperror}) is going to be greater than zero (in the remote scenario where  (\ref{eq:maximization}) is equal to zero). A better initialization would be to set delta to an arbitrary small value, slightly greater than zero.

@@

In case the conditional statement is verified,
the precision tuning phase starts.
For the precision tuning of activation functions, we allow an error that is bounded by \texttt{max\_err}. This is exactly what is described in (\ref{eq:deltaminusmax}).

On the other hand, the quantization of polytopes borders (the region bounds in \statespace) can produce an error that is at most $\varepsilon_{Q}$, in this way we satisfy 
(\ref{eq:quantizationlines}).

In case the conditional fails and \texttt{maxUij} is greater than (or equal to) delta, the controller needs to be re-designed with a robustness value that is at least equal to the current value of \texttt{maxUij}. The same consideration done for the initialization of input parameters holds also here: the constant $\varepsilon_{SAFE}$ is used to relax the value for delta, to a value slightly greater than \texttt{maxUij}. In this way, we aim to give some space to the precision tuning phase in the next iteration of the loop. Otherwise, in case $\varepsilon_{SAFE}$ is equal to zero, the next iteration of the precision tuning phase is going to require unnecessarily wide precision value for p, usually greater than 32 bits. 
We remark that the point of the analysis is to find a low precision configuration for bounds and activation functions: in case the analysis outputs a precision greater than 32bits, we fail in our goal.
Then, with a minimal alteration to the upper bound of delta, we sensibly reduce the precision needed for $F$ and $G$, with respect to the standard 32 bits precision.
